The seasonal cycle of upper-ocean mixing at 8°N in the Bay of Bengal

### 2019-11-19 paper1.tex
Updated figures.
minor edits.

Depth-time maps of the mean squared shear for three shear components \(S\low, S\niwp, S\res\) along with the total shear \(S\tot\) are shown in Figure \ref{fig:shears}.
At all three moorings, energetic shear is observed in January, February, and for an extended period between July and November.
The shear field is relatively weak between mid-March and the beginning of June.
Episodic energetic bursts in near-inertial shear are seen at all three moorings outside these months.
All three moorings see a large rise in low-frequency shear between July and November.
This is an indication of the Sri Lanka Dome and a large anticyclonic eddy through the array as the meandering of the Summer Monsoon Current (note EKE maximum inferred from altimetric data in Figure \ref{fig:spatial}a--e).
The magnitude of the low-frequency shear is comparable to that of near-inertial shear at all three locations during the SW monsoon.
% At NRL4, the low-frequency, near-inertial and residual components contribute equally to total shear.
The residual \(S\res\) is generally weak relative to the \(S\low\) and \(S\niwp\).
The episodic nature of near-inertial shear events prevent a confident estimation of the magnitude of its seasonal cycle given that only one full annual cycle was recorded.
However the seasonal signal in \emph{total} shear — high shear between June and February, and low shear between March and June — is robust and consistent across all three moorings.

@@

A non-dimensional parameter that characterizes the transition from laminar to turbulent flow is the buoyancy Reynolds number \(\Rey_b = ε/(νN^2)\).
When \(ε \approx \SI{1e-9}{\W\per\kg}\), \(N²≈ \SI{1e-3}{\per\s\squared}\) \citep{Jinadasa2016,St.Laurent2017} and molecular viscosity \(ν \approxeq \SI{1e-6}{\m\squared\per\s}\), $\Rey_b \approx 1$.
At such low values of $\Rey_b$, overturning turbulence ceases to exist and total diffusivity asymptotes to \(κ_T\) in direct numerical simulations as well as experiments (for e.g. \citealp{Ivey2008}, their Figure 2; \citealp{Itsweire1993}).
The microstructure $ε$ measurements of \cite{Jinadasa2016} and \cite{St.Laurent2017} then independently indicate that near-molecular diffusivities are frequently present in the Bay.

Low thermocline diffusivities are predicted by the finestructure internal-wave scaling of \cite{Henyey1986} and have been observed previously at low latitudes in the Pacific and Atlantic: \(K_ρ \approx (1\mhyph 3) \times \SI{1e-6}{\m\squared\per\s} (5-15 κ_T)\) for latitudes south of 10°N in \cite{Gregg2003}.
Our lowest observed values during March, April at approximately \SIrange{80}{100}{m} depths are frequently lower than those observations (Figures \ref{fig:vert} and \ref{fig:nrl}b).
Low values of \(K_T\) are perhaps not surprising given the observations summarized above but these χpod observations are the first to show that extremely low mixing (\(K_T ≤ 1\mhyph 10 κ_T\)) persists for multiple weeks at multiple locations in the south-central Bay (Figures \ref{fig:nrl}b and \ref{fig:vert}).

@@

\begin{figure*}
\centering
\includegraphics[width=39pc]{figure4-nrl-spectra.pdf}
\caption{\label{fig:nrlspectra}
Rotary power spectral density of total shear \(S\tot\) at all three moorings estimated using the multitaper method. (a, c, e) Eulerian estimate at \SI{152}{m}; (b, d, f) isothermal estimate at the \SI{18}{\celsius} isotherm. Lowpass, near-inertial and near-tidal bands (colored shading: gray, green, orange respectively) as well as percentage of total shear variance in each band (colored text) are shown. Vertical lines mark \(f_0\), the diurnal frequency, \(ω_{M2} - f_0\) and \(ω_{M2} + f_0\). Clockwise and counter-clockwise spectra are in black and red respectively.}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figure5-shears.png}
\caption{\label{fig:shears}
Weekly running mean-squared shear for the three moorings (top to bottom): (a-c) total shear \(S\tot^2\); (d-f) low-frequency shear \(S\low^2\); (g-i) total near-inertial shear \(S\niwp^2\) (j-l) residual shear \(S\res^2\). All components are normalized by the normalized by 30-day lowpass filtered \(N^2_T\). χpod depths for both χpods are shown in black in all panels. White contours mark levels \num{2e-3}, \num{4e-3} and \num{6e-3}.}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figure6-nrl5.pdf}
\caption{\label{fig:nrl}
A year of observations at NRL5, 105m. Time series of daily averaged quantities: (a) Tropflux wind stress; (b) daily averaged \(K_T\); (c) turbulent heat and salt fluxes \(J_q^t, J_s^t\); (d) Buoyancy frequency \(N^2\) and temperature contribution to \(N^2\), \(g \alpha T_z\); (e) Weekly running mean of filtered squared shear magnitude normalized by $N²$: low pass in black $S\low$, near-inertial bandpass $S\niwp$ in green and the residual $S\res$ in orange; (f) \(\chi\)pod depth. Background colors mark seasons; white region indicates time period shown in Figure \ref{fig:nrl5-niw}.}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figure7-profile.pdf}
\caption{\label{fig:vert}
\small The seasonal cycle of \(K_T\) at 8°N. Vertical profile of hourly averaged \(K_T\) formed by combining all estimates in density bins (Section \ref{sec:profile}). PDFs as well as means and medians are shown. Bins are marked by \(\rho-1000\). Orange horizontal lines mark the climatological depth of the \(S=34.75\) isohaline at 8\(\degree\)N estimated using the Argo climatology. Vertical lines mark the standard deviation of measurement depths in each bin --- these lines tend to overlap each other. Each PDF is colored according to data coverage: one means that there is at least one hourly estimate for every hour in the season.}
\end{figure*}

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figure9-tidal-pumping.pdf}
\caption{\label{fig:nrl5-niw}
An example of pumping of the near-inertial shear layers past the \(\chi\)pod by the \(M_2\) tide at NRL5. The time period of focus is highlighted in white in Figure \ref{fig:nrl}. Time series of (a) turbulent diffusivity \(K_T\), (b) zonal shear and (c) near-inertial and near-tidal shear on two isotherms for a period of high mixing associated with downward propagating near-inertial energy. Horizontal lines indicate the inertial period (3.79 days; labelled \(f_0\)) and the \(M_2\) period (12.42 hours; labelled \(M_2\)). Also shown in (b) are the depth of the \(\chi\)pod and two isotherms (\SI{17}{\celsius}, \SI{19.5}{\celsius}). (c) Near-inertial shear dominates near-tidal shear by an order of magnitude on the two isotherms (\SI{17.0}{\celsius}, \SI{19.5}{\celsius}). The two time series are obtained by first interpolating total shear to isothermal space and then filtering as in Section \ref{sec:results}.}
\end{figure*}

>>>>

Depth-time maps of the mean squared shear for three shear components \(S\low, S\niwp, S\res\) along with the total shear \(S\tot\) are shown in Figure \ref{fig:shears} (normalized by the temperature contribution to stratification $N_T²$).
At all three moorings, energetic shear is observed in January, February, and for an extended period between July and November.
The shear field is relatively weak between mid-March and the beginning of June.
Episodic energetic bursts in near-inertial shear are seen at all three moorings outside these months.
All three moorings see a large rise in low-frequency shear between July and November.
This is an indication of the Sri Lanka Dome and a large anticyclonic eddy through the array as the meandering of the Summer Monsoon Current (note EKE maximum inferred from altimetric data in Figure \ref{fig:spatial}a--e).
The magnitude of the low-frequency shear is comparable to that of near-inertial shear at all three locations during the SW monsoon.
% At NRL4, the low-frequency, near-inertial and residual components contribute equally to total shear.
The residual \(S\res\) is generally weak relative to the \(S\low\) and \(S\niwp\).
The episodic nature of near-inertial shear events prevent a confident estimation of the magnitude of its seasonal cycle given that only one full annual cycle was recorded.
However the seasonal signal in \emph{total} shear — high shear between June and February, and low shear between March and June — is robust and consistent across all three moorings.

@@

A non-dimensional parameter that characterizes the transition from laminar to turbulent flow is the buoyancy Reynolds number \(\Rey_b = ε/(νN^2)\) \citep[for example][]{Itsweire1993}.
When \(ε \approx \SI{1e-9}{\W\per\kg}\), \(N²≈ \SI{1e-3}{\per\s\squared}\) \citep{Jinadasa2016,St.Laurent2017} and molecular viscosity \(ν \approxeq \SI{1e-6}{\m\squared\per\s}\), $\Rey_b \approx 1$.
At such low values of $\Rey_b$, overturning turbulence ceases to exist and total diffusivity asymptotes to \(κ_T\) in direct numerical simulations as well as experiments (for e.g. \citealp{Ivey2008}, their Figure 2; \citealp{Itsweire1993}).
The microstructure $ε$ measurements of \cite{Jinadasa2016} and \cite{St.Laurent2017} then independently indicate that weakly turbulent flows with near-molecular diffusivities are present in the Bay.

Low thermocline diffusivities are predicted by the finestructure internal-wave scaling of \cite{Henyey1986} and have been observed previously at low latitudes in the Pacific and Atlantic: \(K_ρ \approx (1\mhyph 3) \times \SI{1e-6}{\m\squared\per\s} (5-15 κ_T)\) for latitudes south of 10°N in \cite{Gregg2003}.
Our lowest observed values during March, April at approximately \SIrange{80}{100}{m} depths are frequently lower than those observations (Figures \ref{fig:vert} and \ref{fig:nrl}b).
The extended period of low \(K_T\) values is perhaps unsurprising given the observations summarized above and that the transition months of March and April are a period of very low wind energy input i.e. weak inertial inertial shear; weak mean flows i.e. weak low frequency shear; and considerable stratification (note low $S²/N²$ in Figure \ref{fig:shears}).
However these χpod observations are the first to show that extremely low mixing (\(K_T ≤ 1\mhyph 10 κ_T\)) persists for multiple weeks at multiple locations in the south-central Bay (Figures \ref{fig:nrl}b and \ref{fig:vert}).

@@

\begin{figure*}
\centering
\includegraphics[width=33pc]{figure4-nrl-spectra.pdf}
\caption{\label{fig:nrlspectra}
Rotary power spectral density of total shear \(S\tot\) at all three moorings estimated using the multitaper method. (a, c, e) Eulerian estimate at \SI{152}{m}; (b, d, f) isothermal estimate at the \SI{18}{\celsius} isotherm. Lowpass, near-inertial and near-tidal bands (colored shading: gray, green, orange respectively) as well as percentage of total shear variance in each band (colored text) are shown. Vertical lines mark \(f_0\), the diurnal frequency, \(ω_{M2} - f_0\) and \(ω_{M2} + f_0\). Clockwise and counter-clockwise spectra are in black and red respectively.}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figure5-shears.png}
\caption{\label{fig:shears}
Weekly running mean-squared shear for the three moorings (top to bottom): (a-c) total shear \(S\tot^2\); (d-f) low-frequency shear \(S\low^2\); (g-i) total near-inertial shear \(S\niwp^2\) (j-l) residual shear \(S\res^2\). All components are normalized by the normalized by 30-day lowpass filtered \(N_T² = g \alpha T_z\). Regions with $N_T² < \SI{1e-5}{\per\second\squared}$ are excluded. χpod depths for both χpods are shown in black in all panels. White contours mark levels \num{0.75}, \num{1.25}.}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figure6-nrl5.pdf}
\caption{\label{fig:nrl}
A year of observations at NRL5, 105m. Time series of daily averaged quantities: (a) Tropflux wind stress; (b) daily averaged \(K_T\); (c) turbulent heat and salt fluxes \(J_q^t, J_s^t\); (d) Buoyancy frequency \(N^2\) and temperature contribution to \(N^2\), \(N_T² = g \alpha T_z\); (e) Weekly running mean of filtered squared shear magnitude normalized by $N²$: low pass in black $S\low$, near-inertial bandpass $S\niwp$ in green and the residual $S\res$ in orange; (f) \(\chi\)pod depth. Background colors mark seasons; white region indicates time period shown in Figure \ref{fig:nrl5-niw}.}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=33pc]{figure7-profile.pdf}
\caption{\label{fig:vert}
\small The seasonal cycle of \(K_T\) at 8°N. Vertical profile of hourly averaged \(K_T\) formed by combining all estimates in density bins (Section \ref{sec:profile}). PDFs as well as means and medians are shown. Bins are marked by \(\rho-1000\). Orange horizontal lines mark the climatological depth of the \(S=34.75\) isohaline at 8\(\degree\)N estimated using the Argo climatology. Vertical lines mark the standard deviation of measurement depths in each bin --- these lines tend to overlap each other. Each PDF is colored according to data coverage: one means that there is at least one hourly estimate for every hour in the season.}
\end{figure*}

@@

\begin{figure*}
\centering
\includegraphics[width=39pc]{figure9-tidal-pumping.pdf}
\caption{\label{fig:nrl5-niw}
An example of pumping of the near-inertial shear layers past the \(\chi\)pod by the \(M_2\) tide at NRL5. The time period of focus is highlighted in white in Figure \ref{fig:nrl}. Time series of (a) turbulent diffusivity \(K_T\), (b) zonal shear and (c) near-inertial and near-tidal shear on two isotherms for a period of high mixing associated with downward propagating near-inertial energy. Horizontal lines indicate the inertial period (3.79 days; labelled \(f_0\)) and the \(M_2\) period (12.42 hours; labelled \(M_2\)). Also shown in (b) are the depth of the \(\chi\)pod and two isotherms (\SI{17}{\celsius}, \SI{19.5}{\celsius}). (c) Near-inertial shear dominates near-tidal shear by an order of magnitude on the two isotherms (\SI{17.0}{\celsius}, \SI{19.5}{\celsius}). The two time series are obtained by first interpolating total shear to isothermal space and then filtering as in Section \ref{sec:results}.}
\end{figure*}
